Lower Bounds on the Minimax Risk for the \\ Source Localization Problem

### 2017-01-25 paper.tex
Make miscellaneous minor edits.

We give lower bounds on the minimax risk (mean squared-error in location) in
estimating the location of the source, which apply to all estimators, for
\textcolor{blue}{certain classes} of impulse responses of the low-pass filter,
when using a unformly distributed sensor array. We show that
\textcolor{blue}{for sensors of a fixed size}, the lower bound decays with
increasing numbers of sensors. We also perform a preliminary numerical analysis
to understand the effect of increasing the number of sensors to infinity by
shrinking their size, wherein the bound saturates for large sensor numbers.
\textcolor{blue}{In the second scenario, it is seen that there is greater
benefit to increasing the number of sensors} as the signal-to-noise increases.
Our bounds are the first to give a scaling for the minimax risk in terms of the
number of sensors used.

@@

Hence, we believe that this paper is the first to give lower bounds on source
localization error (measured in an $\ell_2$-distance sense) in estimating the
location of a one-dimensional point source, which is observed by sensors
through a diffusive medium (treated as a low-pass filter), and corrupted by
additive white Gaussian noise. We describe the problem setting in detail in
Section~\ref{sec:source-localization} and the minimax techniques and our main
results in Section~\ref{sec:minimax-lower-bounds}. We conclude in
Section~\ref{sec:discussion} with a discussion on how our results can be
extended to other simple settings.

For the application of EEG source localization, it might be possible to find
algorithms which achieve similar scalings as the lower bounds presented here.
It is also hoped that the techniques presented here will help us derive bounds
in more difficult settings, such as spherical and realistic brain models.

@@

All signals on the aforementioned circular domain are of the form
$f:\Theta\mapsto\mathbb{R}$.  The point source located at $\theta$ is
represented by the impulse signal, $f(s;\theta) = \delta(s - \theta)$, where
$\delta(\cdot)$ is the Dirac delta function.  The sensors observe this signal
through the diffusive medium, which, intuitively speaking, blurs the impulse.
More concretely, we assume that the medium is linear and shift-invariant, so
that it can be represented by a spatial impulse response. Then, blurring would
correspond to low-pass filtering.  Let this impulse response be given by
$g(s)$. Then, the noiseless, continuous-space signal (post-filtering and
pre-sampling) is given by the convolution, $x(s; \theta) = (g*f)(s) = g(s -
\theta)$. Here, we further make the simplifying assumption that $g(s)$ has a
sufficiently restricted support, so that aliasing effects are avoided, and
$x(s; \theta)$ is always well-defined. To be precise, $g(s) = 0$ when $\abs{s}
> w/2$, where the ``width'' $w$ of the impulse response satisfies $w < S / 2$
(the reason for the factor of $1/2$ will become clear in a later section). For
the proof, $g(s)$ is also assumed to be Lipschitz continuous with parameter
$\kappa$. This implies that $\abs{g(u) - g(v)} \leq \kappa \abs{u - v} \;
\forall \;\; u, v \in \Theta$.

The sensors sample the continuous-space shifted impulse response, with some
additive noise. We denote the noiseless sampled version of $x(s; \theta)$ by
the $m$-length vector $\v x(\theta)$:
\begin{equation} \label{eq:sampled-signal}
	\v x(\theta) = \bigg[x(0; \theta), \ldots, x\Big(\frac{kS}{m}; \theta\Big), \ldots, x\Big(\frac{(m-1)S}{m}; \theta\Big)\bigg]^T
\end{equation}
where $m$ is the number of sensors and $k \in \{0, 1, \ldots, m-1\}$. The
additive noise is given by $\v \epsilon$ (to be described shortly), and the
noisy samples are represented by
\begin{equation} \label{eq:sensor-obs}
	\v y = \v x(\theta) + \v \epsilon.
\end{equation}
In the source localization problem, we receive $n$ trials\footnote{The word
``trials'' comes from neuroscience literature, where several trials of an
experiment are run with a single participant. The participant is given the same
stimulus in each trial, evoking the same response in the brain (usually with
some trial-to-trial variability). For our purposes, we assume that trials are
effectively i.i.d. realizations of sensor noise, with no trial-to-trial
variability.} (i.e., $n$ realizations) of $\v y$ according to
equation~\eqref{eq:sensor-obs}, for the same point source $\theta$, using which
we need to estimate the location of the source.

The complete signal model described in this section is summarized in
Fig.~\ref{fig:signal-model}.

@@

\textcolor{blue}{We start by lower bounding the maximum risk of the estimation problem with the
average risk of a multiple hypothesis testing problem.} For this, we first need
to define a $2\delta$-packing:%
\begin{definition}
	A set $\Theta_{\mathcal{V}} = \{ \theta_v : v \in \mathcal{V} \}$ for some
	finite index set $\mathcal{V} \subset \mathbb N$ is said to be a
	$2\delta$-packing in the $\rho$-metric if
	\begin{equation}
		\rho(\theta_i, \theta_j) \geq 2\delta \qquad \forall \;\; \theta_i, \theta_j \in \Theta_{\mathcal{V}}.
	\end{equation}
\end{definition}
\begin{theorem} \label{thm:est-to-testing}%
	If we can find a $2\delta$-packing $\Theta_{\mathcal{V}}$ of $\Theta$, then
	we can lower bound the minimax estimation risk by the average testing risk:
	\begin{equation}
		\mathfrak{M}_n(\mathcal{P}, \Phiorho) \geq \Phi(\delta) \inf_\psi \mathbb P (\psi(Y^n) \neq V)
	\end{equation}
	where $V$ is the unknown true hypothesis which takes values uniformly from
	$\mathcal{V}$, and $\psi$ is our estimate of the hypothesis.
\end{theorem}
For a proof of this theorem, we refer the reader to Proposition~13.3
in~\cite{Duchi2015Information}. Intuitively, Theorem~\ref{thm:est-to-testing}
says that the error in estimating $\theta_i$ is likely to be large if it is
difficult to distinguish $\theta_i$ from $\theta_j$, i.e.\ if the probability
of error is high.

@@

Another point on sensor placement bears repeating: the bounds derived in this
paper apply to only a specific sensor configuration, i.e., the uniform one.
Intuitively, this is the correct choice in the minimax setting, since any
non-uniform placement will have one or more points which is far from all
sensors. The ``maximization'' over $\theta \in \Theta$ would then select a
$\theta$ for which the value of the impulse response at the sensors is
smallest, so that the ratio of mean-sensor-value to noise variance (a form of
``SNR'') is minimized.

>>>>

We give lower bounds on the minimax risk (mean squared-error in location) in
estimating the location of the source, which apply to all estimators, for
\textcolor{blue}{certain classes of diffusive media represented by low-pass
filters}, when using a unformly distributed sensor array. We show that
\textcolor{blue}{for sensors of a fixed size}, the lower bound decays with
increasing numbers of sensors. We also perform a preliminary numerical analysis
to understand the effect of increasing the number of sensors to infinity by
shrinking their size, wherein the bound saturates for large sensor numbers.
\textcolor{blue}{In the second scenario, it is seen that there is greater
benefit to increasing the number of sensors} as the signal-to-noise increases.
Our bounds are the first to give a scaling for the minimax risk in terms of the
number of sensors used.

@@

Hence, we believe that this paper is the first to give lower bounds on source
localization error (measured in an $\ell_2$-distance sense) in estimating the
location of a one-dimensional point source, which is observed by sensors
through a diffusive medium (treated as a low-pass filter), and corrupted by
additive white Gaussian noise. We describe the problem setting in detail in
Section~\ref{sec:source-localization} and the minimax techniques and our main
results \textcolor{blue}{(for a naive and then for a more physical sensor
model)} in Section~\ref{sec:minimax-lower-bounds}. We conclude in
Section~\ref{sec:discussion} with a discussion on the goodness of our bounds,
and on how our results can be extended to other simple settings.

\textcolor{blue}{
For EEG source localization, we hope to find algorithms that achieve scalings
close to the bounds we present.  We also expect to build upon the techniques
presented in this paper to derive bounds for spherical and realistic brain
models.
}

@@

%All ``continuous-space'' signals (analogous to continuous- and
	%discrete-\emph{time} signals) on the aforementioned circular domain are
All signals of the form $f:\Theta\mapsto\mathbb{R}$.  The point source located at $\theta$
is represented by the impulse signal, $f(s;\theta) = \delta(s - \theta)$, where
$\delta(\cdot)$ is the Dirac delta function.  The sensors observe this signal
through the diffusive medium, which, intuitively speaking, blurs the impulse.
More concretely, we assume that the medium is linear and shift-invariant, so
that it can be represented by a spatial impulse response. Then, blurring would
correspond to low-pass filtering.  Let this impulse response be given by
$g(s)$. Then, the noiseless, continuous-space signal (post-filtering and
pre-sampling) is given by the convolution, $x(s; \theta) = (g*f)(s) = g(s -
\theta)$. Here, we further make the simplifying assumption that $g(s)$ has a
sufficiently restricted support, so that aliasing effects are avoided, and
$x(s; \theta)$ is always well-defined. To be precise, $g(s) = 0$ when $\abs{s}
> w/2$, where the ``width'' $w$ of the impulse response satisfies $w < S / 2$
(the reason for the factor of $1/2$ will become clear in a later section). For
the proof, $g(s)$ is also assumed to be Lipschitz continuous with parameter
$\kappa$, i.e., $\abs{g(u) - g(v)} \leq \kappa \abs{u - v} \; \forall \;\; u, v
\in \Theta$.

The sensors sample the shifted continuous-space impulse response, with some
additive noise. We denote the noiseless sampled version of $x(s; \theta)$ by
the $m$-length vector $\v x(\theta)$:
\begin{equation} \label{eq:sampled-signal}
	\v x(\theta) = \bigg[x(0; \theta), \ldots, x\Big(\frac{kS}{m}; \theta\Big), \ldots, x\Big(\frac{(m-1)S}{m}; \theta\Big)\bigg]^T
\end{equation}
where $m$ is the number of sensors and $k \in \{0, 1, \ldots, m-1\}$. The
additive noise is given by $\v \epsilon$ (to be described shortly), and the
noisy samples are represented by
\begin{equation} \label{eq:sensor-obs}
	\v y = \v x(\theta) + \v \epsilon.
\end{equation}
In the source localization problem, we receive $n$ trials\footnote{The word
``trials'' comes from neuroscience literature, where several trials of an
experiment are run with a single participant. The participant is given the same
stimulus in each trial, evoking the same response in the brain (usually with
some trial-to-trial variability). For our purposes, we assume that trials are
effectively i.i.d. realizations of sensor noise, with no trial-to-trial
variability.} (i.e., $n$ realizations) of $\v y$ according to
equation~\eqref{eq:sensor-obs}, for the same point source $\theta$, using which
we need to estimate the location of the source.
The complete signal model described in this section is summarized in
Fig.~\ref{fig:signal-model}.

@@

\textcolor{blue}{We start by lower bounding the maximum risk of the estimation problem with the
average risk of a multiple hypothesis testing problem.} For this, we first need
to define a $2\delta$-packing.%
\begin{definition}
	A set $\Theta_{\mathcal{V}} = \{ \theta_v : v \in \mathcal{V} \}$ for some
	finite index set $\mathcal{V} \subset \mathbb N$ is said to be a
	$2\delta$-packing in the $\rho$-metric if
	\begin{equation}
		\rho(\theta_i, \theta_j) \geq 2\delta \qquad \forall \;\; \theta_i, \theta_j \in \Theta_{\mathcal{V}}.
	\end{equation}
\end{definition}
\begin{theorem} \label{thm:est-to-testing}%
	If we can find a $2\delta$-packing $\Theta_{\mathcal{V}}$ of $\Theta$, then
	we can lower bound the minimax estimation risk by the average testing risk:
	\begin{equation}
		\mathfrak{M}_n(\mathcal{P}, \Phiorho) \geq \Phi(\delta) \inf_\psi \mathbb P (\psi(Y^n) \neq V)
	\end{equation}
	where $V$ is the unknown true hypothesis which takes values uniformly from
	$\mathcal{V}$, and $\psi$ is our estimate of the hypothesis.
\end{theorem}
For a proof of this theorem, we refer the reader to Proposition~13.3
in~\cite{Duchi2015Information}. Intuitively, Theorem~\ref{thm:est-to-testing}
says that the error in estimating $\theta_i$ is likely to be large if it is
difficult to distinguish $\theta_i$ from $\theta_j$, i.e.\ if the probability
of error is high.

@@

\textcolor{blue}{
Another point on sensor placement bears repeating: the bounds derived in this
paper apply to only a specific sensor configuration, i.e., the uniform one.
Intuitively, this is the correct choice in the minimax setting, since any
non-uniform placement will have one or more points which is far from all
sensors. The ``maximization'' over $\theta \in \Theta$ would then select a
$\theta$ for which the value of the impulse response at the sensors is
smallest, so that the ratio of mean-sensor-value to noise variance (a form of
``SNR'') is minimized.
}
